---
title: "I did 'Advent of Code' and here is what I learned"
author: "Emil Hvitfeldt"
format:
  revealjs: 
    theme: [default, custom.scss]
---

##  {background-image="images/tree-empty.png" background-size="contain"}

::: {.notes}
Slide 1 / 40
:::

##  {background-image="images/example.png" background-size="contain"}

::: {.notes}
Slide 2 / 40
:::

##  {background-image="images/example-story.png" background-size="contain"}

::: {.notes}
Slide 3 / 40
:::

##  {background-image="images/example-problem.png" background-size="contain"}

::: {.notes}
Slide 4 / 40
:::

##  {background-image="images/example-test.png" background-size="contain"}

::: {.notes}
Slide 5 / 40
:::

##  {background-image="images/one-to-many.png" background-size="contain"}

::: {.notes}
Slide 6 / 40
:::

##  {background-image="images/time.gif" background-size="cover"}

::: {.notes}
Slide 7 / 40
:::

# 2020 December 1st

::: {.notes}
Slide 8 / 40
:::

# Find the two entries that sum to 2020

::: {.notes}
Slide 9 / 40
:::

# My Solution

```{r, echo=TRUE, eval=FALSE}
input <- as.numeric(readLines("01-input"))

sum2 <- function(input, target) {
  for (i in input) for (j in input) if (i + j == target) return(i * j)
}

sum2(input, 2020)
```

::: {.notes}
Slide 10 / 40
:::

# 502 Bad Gateway 🔥🔥🔥

::: {.notes}
Slide 11 / 40
:::

# Possible Approaches

::: incremental
-   loops
-   expand.grid
-   sample
-   Rockstar
:::

::: {.notes}
Slide 12 / 40
:::

------------------------------------------------------------------------

::: {style="font-size:14pt;"}
The goal is to extinguish my desparation\
Knock the goal down

The enemy are accountants\
Knock the enemy down

The system is the enemy\
The machine is the enemy

Listen to the money\
Until the money is gone\
Cast the money into the fire\
Let the rage at the enemy be the fire\
Build the enemy up\
Listen to the money

While the enemy is higher than the system\
Let chaos be the rage at the system\
Let destruction be the rage at the machine\
If the goal is chaos with destruction\
Break it down

Build the machine up\
If the enemy is higher than the machine\
Take it to the top

Build the system up\
The machine is the system

Shout chaos of destruction
:::

::: aside
::: {style="font-size:14pt;"}
https://old.reddit.com/r/adventofcode/comments/k4e4lm/2020_day_1\_solutions/ge8uwqi/
:::
:::

::: {.notes}
Slide 13 / 40
:::

# Types of problems

-   recursive
-   Grid based
-   parsable

::: {.notes}
Slide 14 / 40
:::

## Parsable Solutions

::: columns
::: {.column width="50%"}
```{r, eval=FALSE, echo=TRUE}
forward 2
down 4
down 3
up 4
down 1
down 8
up 9
forward 1
down 9
forward 6
down 7
forward 1
down 1
up 2
forward 8
down 3
down 9
down 3
```
:::

::: {.column width="50%"}
:::
:::

::: {.notes}
Slide 15 / 40
:::

## Parsable Solutions

::: columns
::: {.column width="50%"}
```{r, eval=FALSE, echo=TRUE}
forward(2)
down(4)
down(3)
up(4)
down(1)
down(8)
up(9)
forward(1)
down(9)
forward(6)
down(7)
forward(1)
down(1)
up(2)
forward(8)
down(3)
down(9)
down(3)
```
:::

::: {.column width="50%"}
-   modify the input

::: incremental
-   define `forward()`, `down()` and `up()`
-   run `eval(parse(text = input))`
-   ...
-   profit
:::
:::
:::

::: {.notes}
Slide 16 / 40
:::

# memory exhausted (limit reached)

::: {.notes}
Slide 17 / 40
:::

## Lanternfish

Initial state: 3,4,3,1,2\

::: {.notes}
Slide 18 / 40
:::

## Lanternfish

Initial state: 3,4,3,1,2\
After  1 day:  2,3,2,[0]{.green},1\

::: {.notes}
Slide 19 / 40
:::

## Lanternfish

Initial state: 3,4,3,1,2\
After  1 day:  2,3,2,[0]{.green},1\
After  2 days: 1,2,1,[6]{.darkgreen},0,[8]{.glow}\

::: {.notes}
Slide 20 / 40
:::

## Lanternfish

Initial state: 3,4,3,1,2\
After  1 day:  2,3,2,[0]{.green},1\
After  2 days: 1,2,1,[6]{.darkgreen},0,[8]{.glow}\
After  3 days: [0]{.green},1,[0]{.green},5,6,7,[8]{.glow}\
After  4 days: [6]{.darkgreen},[0]{.green},[6]{.darkgreen},4,5,6,7,[8]{.glow},[8]{.glow}\
After  5 days: 5,[6]{.darkgreen},5,3,4,5,6,7,7,[8]{.glow}\
After  6 days: 4,5,4,2,3,4,5,6,6,7\
After  7 days: 3,4,3,1,2,3,4,5,5,6\
After  8 days: 2,3,2,[0]{.green},1,2,3,4,4,5\
After  9 days: 1,2,1,[6]{.darkgreen},[0]{.green},1,2,3,3,4,[8]{.glow}\
After 10 days: [0]{.green},1,0,5,[6]{.darkgreen},0,1,2,2,3,7,[8]{.glow}

::: {.notes}
Slide 21 / 40
:::

## Lanternfish

### Part 1

[How many lanternfish would there be after 80 days?]{.glow}

::: {.notes}
Slide 22 / 40
:::

## Lanternfish

### Part 1

How many lanternfish would there be after 80 days?

### Part 2

[How many lanternfish would there be after 256 days?]{.glow}

::: {.notes}
Slide 23 / 40
:::

## Lanternfish

### Part 1

How many lanternfish would there be after 80 days? [\~1.5 MB]{.glow}

### Part 2

How many lanternfish would there be after 256 days?

::: {.notes}
Slide 24 / 40
:::

## Lanternfish

### Part 1

How many lanternfish would there be after 80 days? [\~1.5 MB]{.glow}

### Part 2

How many lanternfish would there be after 256 days? [\~6.58 TB]{.glow}

::: {.notes}
Slide 25 / 40
:::

## General workflow

::: columns
::: {.column width="50%"}
-   [reading in data]{.glow}

-   doing calculations

-   unit tests

-   Solution
:::

::: {.column width="50%"}
- {readr}
- `data.table::fread()`
- {jsonlite}
- `read_parquet()`
:::

:::

::: {.notes}
Slide 26 / 40
:::

## General workflow

::: columns
::: {.column width="50%"}
-   reading in data

-   [doing calculations]{.glow}

-   unit tests

-   Solution
:::

::: {.column width="50%"}
- vectors
- matrices
- data.frames
- lists
:::

:::

::: {.notes}
Slide 27 / 40
:::

## General workflow

-   reading in data

-   doing calculations

-   [unit tests]{.glow}

-   Solution

::: {.notes}
Slide 28 / 40
:::

## General workflow

-   reading in data

-   doing calculations

-   unit tests

-   [Solution]{.glow}

::: {.notes}
Slide 29 / 40
:::

# You might learn something

::: {.notes}
Slide 30 / 40
:::

---

## Ways you can play this game

::: incremental
-   You can go for speed
-   shortest solution
-   no dependencies
-   try another language
:::

::: {.notes}
Slide 31 / 40
:::

---

##  {background-image="images/speed.gif" background-size="cover"}

::: {.notes}
Slide 32 / 40
:::

---

##  {background-image="images/drob-dominate.png" background-size="contain"}

::: {.notes}
Slide 33 / 40
:::

# go at your own phase

::: {.notes}
Slide 34 / 40
:::

##  {background-image="images/private-dashboard.png" background-size="contain"}

::: {.notes}
Slide 35 / 40
:::

##  {background-image="images/tanho.png" background-size="contain"}

::: {.notes}
Slide 36 / 40
:::

# The R community is awesome!

::: {.notes}
Slide 37 / 40
:::

---

## Many different problems

::: incremental
- Squares With Three Sides (2016 day 3)
- Encoding Error (2020 day 9)
- Wizard Simulator 20XX (2015 day 22)
:::

::: {.notes}
Slide 38 / 40

Encoding error: first number in the list (after the preamble) which is not the sum of two of the 25 numbers before it.
:::

---

## the problems are CLEARLY written

::: {.notes}
Slide 39 / 40
:::

---

# Take home

::: incremental
- You might learn something
- Join the Community
- Have fun!
:::

::: {.notes}
Slide 40 / 40
:::
